---
title: "Dynamic Copula based portfolio optimization: India"
output: html_document
date: "2024-11-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
This report presents a small empirical analysis of daily returns for ten Indian assets: RELIANCE.NS, TCS.NS, INFY.NS, HDFCBANK.NS, ICICIBANK.NS, HINDUNILVR.NS, ITC.NS, KOTAKBANK.NS, LT.NS, and SBIN.NS.

The dataset spans January 2022 to January 2024, and the analysis uses a 470-day rolling window for return calculations. This window size balances computational efficiency and reduce running time. However, the window size can be adjusted based on specific analysis requirements.
```{r cars}
# Clear environment
rm(list = ls())
gc()

## load necessary libraries
library(copula)
library(sgt)
library(quantmod)
library(PerformanceAnalytics)## for var and cvar
library(quadprog)
library(zoo)
library(doParallel)
library(parallel)
library(foreach)
library(corrplot)
library(ggplot2)
library(reshape2)
library(ggpubr)
library(ADGofTest)## for ad.test

#track running time
strt=Sys.time()

# Set up parallel backend
num_cores <- detectCores() - 1  # Leave one core free
cl <- makeCluster(num_cores)
registerDoParallel(cl)

# Fetch historical stock prices
asset_names <- c("RELIANCE.NS", "TCS.NS", "INFY.NS", "HDFCBANK.NS", "ICICIBANK.NS",
                 "HINDUNILVR.NS", "ITC.NS", "KOTAKBANK.NS", "LT.NS", "SBIN.NS")

getSymbols(asset_names, from = "2022-01-01", to = "2024-01-01")

# Initialize an empty list to store returns
returns_list <- list()

# Loop through each asset to calculate log returns
for (ticker in asset_names) {
  returns_list[[ticker]] <- na.omit(100 * diff(log(Ad(get(ticker)))))
}

# Combine all returns into a matrix
returns_matrix <- coredata(do.call(merge, c(returns_list, all = FALSE)))
colnames(returns_matrix) <- asset_names

### some values which user can change as per the need
window_size <-470
##sample size m
m=5*10^5

d=length(asset_names)
num_windows <- nrow(returns_matrix) - window_size+1


##***++++==correlation heat map to see the portfolio dependence structure&&&***+++
# Correlation matrix
cor_matrix <- cor(returns_matrix)
# Plot the correlation matrix as a heatmap
corrplot(cor_matrix, method = "color", tl.cex = 0.7, addCoef.col = "black", number.cex = 0.5,  )

###*** Define the fitting marginal sgt function&&**$$$$
###*** Fit empirical checkerboard copula&&***
fit.empirical_checkerboard_copula <- function(log_returns_matrix) {
  # Preallocate uniform marginals
  uniform_marginals <- matrix(NA, nrow = nrow(log_returns_matrix), 
                              ncol = d)
  
  # Fit marginals in parallel
  fit_results <-  foreach(i = 1:d, .combine = 'cbind', 
                          .packages = 'sgt') %dopar% {
                            # Try to fit the model and catch errors/warnings
                            pp=log_returns_matrix[, i]
                            result <- tryCatch({
                              # Fit the marginal model
                              sgt_fit <- sgt.mle(X.f = ~as.numeric(pp),
                                                 start = list(mu = mean(pp), 
                                                              sigma = sd(pp), 
                                                              lambda = 0, p = 2, q = 2),
                                                 mean.cent = TRUE, var.adj = TRUE)
                              est=as.numeric(sgt_fit$estimate)
                              list(estimate = est, warnings = NULL)  # No warning
                            }, warning = function(w) {
                              list(estimate = est, warnings = conditionMessage(w))  # Return warning
                            }, error = function(e) {
                              list(estimate = rep(NA,5), warnings = conditionMessage(e))  # Handle error with NA
                            })
                            
                            return(result)  # Return results for each asset
                          }
  
  # Combine estimates and warnings into separate lists
  parest <- do.call(rbind, fit_results[1,])
  warnings_sgtfit <- unlist(fit_results[2,])
  
  
  # Populate uniform_marginals using the fit results
  p_val=numeric(d)
  for (i in 1:d) {
    params <- parest[i,]
    uniform_marginals[, i] <- psgt(log_returns_matrix[, i], 
                                   mu = params[1], sigma = params[2], 
                                   lambda = params[3], p = params[4], 
                                   q = params[5])
    ad_test <- ad.test(log_returns_matrix[, i], 
                       function(x) psgt(x, 
                                        mu = params[1], sigma = params[2], 
                                        lambda = params[3], p = params[4], 
                                        q = params[5]))
    
    p_val[i]=ad_test$p.value
  }
  
  # Fit the empirical checkerboard copula
  copula_fit1 <- empCopula(uniform_marginals,smoothing="checkerboard")
  return(list(copula_fit1,p_val,parest,warnings_sgtfit))
}


####*** Portfolio optimization&&&****
portfolio.opt <- function(simulated_returns,rolling_window,
                          next_window) {
  cov_matrix <- cov(simulated_returns)
  actual_returns <- colMeans(rolling_window)
  
  # Set up quadratic programming problem
  Dmat <- cov_matrix
  dvec <- rep(0, ncol(simulated_returns))
  
  # Constraints: weights sum to 1 and expected return non-negative 
  ## optimal weight portfolio > equal weight portfolio
  Amat <- cbind(1, actual_returns, diag(ncol(simulated_returns)))
  bvec <- c(1, mean(actual_returns), rep(0, ncol(simulated_returns)))
  
  # Solve quadratic programming problem
  opt_result <- tryCatch({
    # Attempt to solve the quadratic programming problem
    result <- solve.QP(Dmat, dvec, Amat, bvec, meq = 1)
    
    # Return the result if no error occurs
    list(result = result, warning = NULL)
    
  }, error = function(e) {
    # Catch and store the error message
    message("Error in portfolio optimization: ", conditionMessage(e))
    list(result = NULL, warning = conditionMessage(e))  # Store the error message as a warning
  }, warning = function(w) {
    # Catch and store the warning message
    message("Warning in portfolio optimization: ", conditionMessage(w))
    list(result = result, warning = conditionMessage(w))  # Store the warning message
  })
  
  
  weights <- opt_result$result$solution
  portfolio_return_nextday <- sum(weights * colMeans(next_window))
  portfolio_variance_nextday <- t(weights) %*% 
    cov(next_window) %*% weights
  
  # Sharpe Ratio (assuming risk-free rate is 0)
  sharpe_ratio <- portfolio_return_nextday / sqrt(portfolio_variance_nextday)
  
  return(list(weights = weights, 
              return = portfolio_return_nextday, 
              sharpe_ratio = sharpe_ratio,
              opt_warning=opt_result$warning))
}


####**** Function to process each rolling window&&***
rollapply_func <- function(rolling_window, next_window) {
  copula_fit <- fit.empirical_checkerboard_copula(rolling_window)
  
  # Simulate returns from the copula
  n_simulations <- m
  simulated_copula <- rCopula(n_simulations, copula_fit[[1]])
  
  # Transform back to original marginals
  simulated_returns <- sapply(1:d, function(i) {
    params <- copula_fit[[3]][i,]
    qsgt(simulated_copula[, i], mu = params[1], sigma = params[2], 
         lambda = params[3], p = params[4], q = params[5])
  })
  max_eigen=max(eigen(cov(rolling_window))$values)
  
  
  # Optimize portfolio
  optimized_portfolio <- portfolio.opt(simulated_returns,
                                       rolling_window, next_window)
  
  # Calculate average return at t+1 i.e. equal weight return for historical data
  avg_return_t1 <- mean(colMeans(next_window))
  avg_return_var_t1 <- t(rep((1/d),d)) %*% cov(next_window) %*% rep((1/d),d)
  sharpe_ratio_t1 <-  avg_return_t1 / sqrt(avg_return_var_t1)
  
  # Combine results for optimal portfolio return at t+1
  return(list(returnweight=c(optimized_portfolio$weights), 
              optfolio=c(optimized_portfolio$return, 
                         optimized_portfolio$sharpe_ratio), 
              eqfolio=c(avg_return_t1,sharpe_ratio_t1),maxeigen=max_eigen,pval= 
                c(copula_fit[[2]]),warningg=c(copula_fit[[4]],
                                              optimized_portfolio$opt_warning)))
}

####**** Run rollapply with parallel processing&&&****
rolling_results <- foreach(i = 1:(num_windows-1),.combine ='cbind' , 
                           .packages = c("zoo", "quantmod","foreach",
                                         "copula", "sgt", "quadprog",
                                         "ADGofTest"))%dopar% {
                                           # Rolling window and t+1 window
                                           roll_window <- returns_matrix[i:(i + window_size - 1), ]
                                           next_row <- returns_matrix[(i+1):(i + window_size), ]                          
                                           result <- tryCatch({
                                             # Assume rollapply_func returns a list containing both numeric and character values
                                             roll_result <- rollapply_func(roll_window, next_row)
                                             
                                             # Extract numeric values (adjust according to the actual structure of roll_result)
                                             numeric_result <- c(roll_result$returnweight, roll_result$optfolio,
                                                                 roll_result$eqfolio, 
                                                                 roll_result$maxeigen, roll_result$pval)
                                             
                                             # Extract any warnings or character values (e.g., roll_result$warning contains warnings)
                                             if(is.null(roll_result$warningg)==F) warning_result <- c(roll_result$warningg)
                                             if(is.null(roll_result$warningg)==T) warning_result <- NULL#"no warning"
                                             # Combine numeric and character results into a single vector
                                             combined_result <- list(numeric_result, warning_result)
                                             
                                           }, error = function(e) {
                                             # In case of an error, return NA for numeric and a generic warning message
                                             message("Error in iteration ", i, ": ", conditionMessage(e))
                                             return(combined_result)
                                           })
                                           
                                           return(result)
                                         }

# Stop the cluster after computation
stopCluster(cl)
Sys.time() - strt
numeric_matrix_out <- do.call(cbind, rolling_results[1,])

# Extract character vectors and combine them into a matrix
character_matrix_warning <- do.call(rbind, rolling_results[2,])

####**** plot the optimal and equal weigted return and sharpe ratio&&***
# Extract the 11th and 14th columns for optimal and equal weighted portfolio returns
optimal_portfolio_return <- t(numeric_matrix_out)[, (d +1)]  # Adjusted for column index
equal_weighted_portfolio_return <- t(numeric_matrix_out)[, (d +3)]  # Adjusted for column index

# Extract the 12th and 16th columns for optimal and equal weighted sharpe ratios
optimal_portfolio_sr <- t(numeric_matrix_out)[, (d +2)]  # Adjusted for column index
equal_weighted_portfolio_sr <- t(numeric_matrix_out)[, (d +4)]  # Adjusted for column index


# Create a data frame for plotting returns
df_returns <- data.frame(
  Time = 1:length(optimal_portfolio_return),
  Optimal = optimal_portfolio_return,
  Equal_Weighted = equal_weighted_portfolio_return
)

df_returns2 <- data.frame(
  Time = 1:length(optimal_portfolio_sr),
  Optimal = optimal_portfolio_sr,
  Equal_Weighted = equal_weighted_portfolio_sr 
)

# Plot 1: Line chart for optimal portfolio return vs equal weighted portfolio return
pl11= ggplot(df_returns, aes(x = Time)) +
  geom_line(aes(y = Optimal, color = "Optimal Portfolio"), size = .5) +
  geom_line(aes(y = Equal_Weighted, color = "Equal Weighted Portfolio"), size = .08) +
  labs(title = "Comparison of Nextday Optimal Portfolio with Other Methods",
       subtitle = "(i) Returns",
       x = NULL,
       y = "Return") +
  scale_color_manual(values = c("Optimal Portfolio" = "blue", "Equal Weighted Portfolio" = "red")) +
  theme_minimal(base_size = 8) +
  theme(legend.title = element_blank(),axis.title = element_text(size = 8),  # Axis titles
        axis.text = element_text(size = 8),   # Axis text
        plot.title = element_text(size = 10),
        legend.position = "none" )

# Plot 2: Line chart for optimal portfolio sharp ratio vs equal weighted portfolio return
pl1= ggplot(df_returns2, aes(x = Time)) +
  geom_line(aes(y = Optimal, color = "Optimal Portfolio"), size = .5) +
  geom_line(aes(y = Equal_Weighted, color = "Equal Weighted Portfolio"), size = .08) +
  labs(subtitle = "(ii) Sharpe Ratio",
       x = "Time_window",
       y = "Sharpe Ratio") +
  scale_color_manual(values = c("Optimal Portfolio" = "blue", "Equal Weighted Portfolio" = "red")) +
  theme_minimal(base_size = 8) +
  theme(legend.title = element_blank(),axis.title = element_text(size = 8),  # Axis titles
        axis.text = element_text(size = 8),   # Axis text
        plot.title = element_text(size = 10),
        legend.position = "none" )
gg1=ggarrange(pl11,pl1,nrow=2,labels=c("C",""),
              font.label = list(size = 9),
              common.legend = T,
              legend="bottom") 
gg1
####****plot the optimal weights and the rolling mean and sd***
optimal_weights=t(numeric_matrix_out)[, (1:d)]
# Create a data frame for plotting optimal weights
df_weights <- data.frame(Time_window = 1:nrow(optimal_weights), 
                         optimal_weights)

# Melt the data to long format for ggplot
df_weights_long <- reshape2::melt(df_weights, id.vars = "Time_window",
                                  variable.name = "Asset", value.name = "Weight")

# Assign asset names to the "Asset" column
df_weights_long$Asset <- factor(df_weights_long$Asset, 
                                labels = asset_names)

# Plot 2: Line chart for optimal asset weights
pl2= ggplot(df_weights_long, aes(x = Time_window, y = Weight, 
                                 color = Asset)) +
  geom_line(size = .5) +
  labs(title = "Rolling Estimated Optimal Weights ",
       x = "Time_window",
       y = "Weight") +
  theme_minimal(base_size = 8) +
  theme(legend.title = element_blank(),axis.title = element_text(size = 8),  # Axis titles
        axis.text = element_text(size = 8),   # Axis text
        plot.title = element_text(size = 10),
        legend.position = "bottom" )+
  geom_hline(yintercept = 1/ncol(returns_matrix), linetype = "dashed", color = "black")  # Add horizontal line

# Function to compute rolling returns and standard deviations
rollapply_func <- function(rolling_window) {
  stock_returns <- colMeans(rolling_window)   # Mean return for each stock
  stock_sd <- apply(rolling_window, 2, sd)    # Standard deviation of each stock
  return(c(stock_returns, stock_sd))
}

# Use rollapply for rolling calculations
rolling_resultss <- rollapply(returns_matrix, width = window_size,
                              FUN = rollapply_func, by.column = FALSE,
                              align = "right")

# Extract the rolling returns and standard deviations from the results
rolling_returns <- rolling_resultss[, 1:d]  # First columns are rolling returns of each stock
rolling_sd <- rolling_resultss[, (d + 1):(2 * d)]  # Next columns are rolling SDs

# (1) Plot Rolling Returns for selected stocks
df_returns <- data.frame(Time = c(seq(0:(num_windows-1))),
                         rolling_returns)
colnames(df_returns)[-1] <- asset_names

p1 <- ggplot(melt(df_returns, id.vars = "Time"),
             aes(x = Time, y = value, color = variable)) +
  geom_line(size = .5) +
  labs(title="Performance Plots",
       subtitle=" (i) Empirical Rolling Average") +
  xlab(NULL) +
  ylab("Rolling Mean") +
  theme_minimal(base_size = 8) +  # Adjust base font size
  theme(legend.position = "none",legend.title = element_blank(),  # Hide legend here
        axis.title = element_text(size = 8),  # Axis titles
        axis.text = element_text(size = 8),   # Axis text
        plot.title = element_text(size = 10)) # Title

# (2) Plot Rolling Standard Deviations for selected stocks
df_sd <- data.frame(Time = c(seq(0:(num_windows-1))),
                    rolling_sd)
colnames(df_sd)[-1] <- asset_names

p2 <- ggplot(melt(df_sd, id.vars = "Time"), 
             aes(x = Time, y = value, color = variable)) +
  geom_line(size = .5) +
  labs(title="",subtitle="(ii) Standard Deviations (SD)") +
  xlab(NULL) +
  ylab("Rolling SD") +
  theme_minimal(base_size = 8) +  # Adjust base font size
  theme(legend.position = "none", legend.title = element_blank(), # Hide legend here
        axis.title = element_text(size = 8),  # Axis titles
        axis.text = element_text(size = 8),   # Axis text
        plot.title = element_text(size = 10)) # Title

p3=   ggarrange(
  p1,                             # First row with one plot
  p2,    # Second row with two plots
  ncol = 2,labels=c("A",""),
  font.label = list(size = 9)
  # Add a common legend
  # Position the legend at the bottom
)

gg2=  ggarrange(p3,pl2,nrow=2,common.legend = T,
                labels=c("","B"),
                font.label = list(size = 9),
                legend="bottom")

gg2

###*** avg of all rolling windows ad.test p-value of sgt fit***
p_valuee=colMeans(t(numeric_matrix_out)[,-c(1:(d+5))])
p_valuee

## total run time
Sys.time()-strt

```


